---
output: github_document
---

```{r, echo=FALSE}
knitr::opts_chunk$set(
  collapse = T, 
  fig.path = "man/figures/README-"
)
```

# `ValleyBikes`

The `ValleyBikes` package aims to make it easier to explore and analyze public data from the Pioneer Valley bikeshare initiative: [Valley Bike](https://valleybike.org/).

## Installation Instructions

The `ValleyBikes` package is online available for everyone. To install you will need the `devtools` package. 

```{r, eval=FALSE}
# install the development version from GitHub
devtools::install_github("Amherst-Statistics/ValleyBikes")
```

To load the package: 

```{r, message=FALSE}
library(ValleyBikes)
library(dplyr)
```

Now we can begin exploring bike data!

## Looking at the Data

There are two tables in our data package: 

- `stations`: contains information about all the stations.
- `routes`: contains the start and end entries for all bike rides (routes) taken.

Let's take a look at the `stations` table: 

```{r}
head(stations)
```

Now let's take a look at the `routes` table: 

```{r}
head(routes)
```

## Analyzing Usage Patterns

Suppose that we want to explore the difference between weekdays and weekends for 2019. We can easily create a plot:

```{r, message = F}
library(ggplot2)
library(lubridate)
```

```{r}
routes %>%
    filter(year(date) >= 2019) %>% 
    mutate(hour = hour(date), day = wday(date, label=T, abbr=T),
           month = month(date, label = T, abbr = T)) %>%
    mutate(weekday = ifelse(day == "Sun" | day == "Mon", "Yes", "No")) %>% 
    count(route_id, hour, weekday, month) %>%
    count(hour, weekday, month) %>% 
    ggplot(aes(x = hour, y = n, color = weekday)) +
    geom_smooth(method = "loess") +
    facet_wrap(~ month) +
    labs(x = "Hour of Day", y = "# of Rides", color = "Weekday?", 
         title = "Valley Bike Usage", subtitle = "Weekdays vs Weekends")
```

From the figure we can see that the gap between weekdays and weekends increases in summer and suddenly dissapears when school starts. 

