---
title: "New Downloading Interface"
output: pdf_document
---

The data is stored in individually compressed files at https://nhorton.people.amherst.edu/valleybikes/ .

```{r}
library(readr)
```

These is how we would read with readr.

```{r}
base_dir <- "https://nhorton.people.amherst.edu/valleybikes/"

Day <- read_csv(paste0(base_dir, "VB_Routes_Data_2018_06_30.csv.gz"),
         skip = 2,
         col_types = cols(
             `Route ID` = col_character(),
             Bike = col_number(),
             Date = col_datetime(format = ""),
             Latitude = col_double(),
             Longitude = col_double(),
             `User ID` = col_character())
         ) %>% janitor::clean_names()
```

However this isn't flexible because it loads the file into memory, while we want to download it.

Here is a function that takes a path to a resource and an output directory and only downloads the file if it hasn't been downloaded yet.

```{r}
# download a single file. this function should contain check if the file was already downloaded.
fetch_file <- function(file_path, output_dir, replace = FALSE) {
    file_name <- basename(file_path)
    output_path <- file.path(output_dir, file_name)
    if(!file.exists(output_path) || replace) {
        download.file(file_path, destfile = output_path)
    }
    return(output_path)
}
```

We want to do this for multiple files through a single function that asks to download all the files within a given set of year/months.

First we create function that downloads a given date, e.g. `2019_07_26`

```{r}
fetch_date <- function(date, output_dir, replace = FALSE) {
    root <- "https://nhorton.people.amherst.edu/valleybikes"
    prefix <- "VB_Routes_Data_"
    suffix <- ".csv.gz"
    file_name <- paste0(prefix, date, suffix)
    file_path <- file.path(root, file_name)
    output_path <- fetch_file(file_path, output_dir, replace)
    return(output_path)
}
```

Now we create a function that downloads all dates between a start and an end date. 

```{r}
fetch_dates <- function(start_date, end_date, output_dir, replace = FALSE) {
    dates <- seq(from = lubridate::as_date(start_date),
                 to = lubridate::as_date(end_date),
                 by = "days")
    dates <- dates %>%
        purrr::map_chr(~ stringr::str_replace_all(.x, "-", "_"))
    output_paths <- dates %>%
        purrr::map_chr(~ fetch_date(.x, output_dir, replace))
    return(output_paths)
}
```
