---
title: "vbvignette"
author: "Emily Lachtara and Shukry Zablah"
date: "October 2019"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vbvignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
#The ValleyBikes Package: Exploring the Pioneer Valley's Electronic Bike System
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(dplyr)
```

## Installing the ValleyBikes Package
To install the package, we use the "library" function. To get an idea of what is included in the package, we can call upon the "help" function.
```{r}
library(ValleyBikes)
```

### It includes

*Data set "routes" with variables: route_id, bike, date, latitude, longitude, user_id
    
*Data set "stations" with variables: Serial_num, Address, Station_name, Num_docks, Latitude, Longitude
    
*More information can be obtained with help(package= "ValleyBikes)

```{r}
help(package= "ValleyBikes")
```

## Exploring the Data

A few natural questions arise from the datasets provided. 

1) Where are these stations located? Can we visualize this?

Using the "leaflet" function in tandem with the stations data, we can create a nice map of the stations.
```{r}
library(leaflet)
stations%>%
bikemap<- leaflet()%>%
  addTiles()%>%
  addMarkers(lng='longitude', lat='latitude')
bikemap


```

2) 
