---
title: "vbvignette"
author: "Emily Lachtara and Shukry Zablah"
date: "October 2019"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{vbvignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
#The ValleyBikes Package: Exploring the Pioneer Valley's Electronic Bike System
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(dplyr)
```

## Installing the ValleyBikes Package
To install the package, we use the "library" function. To get an idea of what is included in the package, we can call upon the "help" function.
```{r}
library(ValleyBikes)
```

### It includes

Data set "routes" with variables: route_id, bike, date, latitude, longitude, user_id

```{r}
head(routes)
```
    
Data set "stations" with variables: serial_num, address, station_name, num_docks, latitude, longitude
   
```{r}
head(stations)
``` 

More information can be obtained with help(package= "ValleyBikes)


### Exploring the Data

A few natural questions arise from the datasets provided. 

## 1) Where are these stations located? Can we visualize this?

Using the "leaflet" function in tandem with the stations data, we can create a nice map of the stations.

```{r}
library(leaflet)

bikemap<-leaflet(stations) %>% addTiles()
bikemap %>% addMarkers(lng=~longitude, lat=~latitude, popup= ~station_name, data=stations)
```

## 2) How does the usage vary across the year?

Using the "lubridate" package we can work with the routes data to visualize variation in usage.

Here we clean up the variable 'date' to fit the ymd_hms format.

```{r}
library(lubridate)

routes2<-routes%>%
   mutate(dates=ymd_hms(date))%>%
   mutate(year=year(dates))%>%
   mutate(month = month(dates, label = T, abbr = T))%>%
   mutate(hour=hour(dates))
```


# Monthly Usage

August sees the most bike usage, whereas there is no usage in the winter months.

```{r, warning=FALSE}
library(ggplot2)
routes2%>%
  group_by(month)%>%
  count()%>%
  filter(month!= "NA")%>%
ggplot(aes(x=month, y=n))+
  geom_bar(stat="identity")+
  labs(title="Monthly Usage" )
```

# Yearly Usage

```{r}
library(xtable)
library(knitr)
kable(routes2%>%
  filter(year>=2018)%>%
  group_by(year)%>%
  summarize(
      number = n()
  ), caption= "Rides per year")
```

# Hourly Usage

The plot of hourly usage follows a nice pattern. However we know that the most popular hour is not 21 (9 pm). So we employ a 'force_tz' to change the time zone to 'Eastern'. We can see that the most popular trip time is 4 pm.

```{r}
routes2%>%
   force_tz(hour, tzone = "US/Eastern")%>%
  group_by(hour)%>%
  count()%>%
ggplot(aes(x=hour, y=n))+
  geom_point()+
  geom_line()+
  ggtitle("Hourly Usage Eastern")
```






