---
title: "Routes"
date: "`r Sys.Date()`"
author: "Shukry Zablah and Emily Lachtara"
output: pdf_document
---

## Setup

```{r}
library(dplyr)
library(tidyr)
library(vroom)
library(lubridate)
library(ggplot2)
```

## Ingestion

We import the slim version of the data that contains only the start and end entries for any given route, throughout all available data. 

```{r}
Routes <- vroom("../data/slim.tsv",
                col_types = cols(
                    route_id = col_character(),
                    bike = col_number(),
                    date = col_datetime(format = ""),
                    latitude = col_double(),
                    longitude = col_double(),
                    user_id = col_character()
                ))

Routes %>% head()
```

## Cleaning

Two things.

1) Our function to slim down the data actually took care of badly downloaded data so will get some lines that are only NAs.
2) The data actually has duplicate entries.

```{r}
Routes <- Routes %>%
    drop_na(-date)  %>% # 9 entries are NA only
  #  drop_na() %>%  # around 400 entries have no date info
    distinct()
```

We removed a tiny amount of data but it will simplify our lives to have data without NAs.

## Exploratory Data Analysis and Further Cleanup

###  We have data for what days? 

We know Valley Bike has only been a thing since 2017. Let's look at the data. 

```{r}
Routes %>%
    mutate(date=round_date(date, unit="day")) %>%
    count(date)%>%
    ggplot(aes(x=date, y=n)) + geom_point() + geom_smooth()
       
```

Let's take a closer look. 

```{r}
Routes %>%
    arrange(date) %>%
    head(15)
```

Fix up the plot by filtering.

```{r}
Routes%>%
  mutate(date=round_date(date, unit="day"))%>%
  filter(date >= "2017-01-01")%>%
    count(date)%>%
    ggplot(aes(x=date, y=n)) + geom_point() + geom_smooth(span=0.2)
```

There is a bimodal distribution of bike usage, where the peak of the usage is in the summer of 2018 and 2019. There does not appear to be much usage in the winter months.

## 2) How many unique routes are there? 
```{r}
Routes %>%
  filter(latitude>42)%>%
  count(route_id)%>%
  count()

```

## 3) Where are the stations? 
```{r}
Routes %>%
  filter(latitude>42)%>%
  count(route_id)%>%
  count()

```
